Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-10-28 grove-paper.tex
Commit yesterday's changes

%% Title information
\title[Grove]{Grove: A Convergent Collaborative Structure-Editor~Calculus}
%\subtitle{Subtitle}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
\autoref{sub:Single User Edits} explores edits by a single user, Alice,
and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

TODO: move subcaptions into master caption
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{A simple tree with a hole.
\autoref{fig:A simple tree with a hole-a}:~User interface.
\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
\label{fig:A simple tree with a hole}
\end{figure}

@@

Program state is a directed graph with a distinguished root vertex.
%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole-b}, \Vertex{2} and \Vertex{4} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

@@

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{2}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

@@

\subsection{Single User Edits}
\label{sub:Single User Edits}

In the remainder of this section,
we will consider user edits, starting
in this subsection with single-user
edits and continuing in \autoref{sub:Multi-user Edits}
with multi-user edits.

Each user edit that we consider corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

User edits act relative to a cursor. 
We do not model cursor sharing and the cursor does not appear in the underlying graph.
We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\columnwidth}
\centering
\caption{}
\label{fig:Construction-a}
TODO: remove this figure
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\columnwidth}
\centering
\caption{}
\label{fig:Construction-b}
\newedge{ConstructionY}{R}%
\newvertex{ConstructionY}{y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vConstructionY" [> "\eConstructionY"]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Construction.
\autoref{fig:Construction-a}:~User interface.
\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{6} and \Edge{5}.}
\label{fig:Construction}
\end{figure}

\todo{"user edits" and "graph edits"}
To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole} and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This edit corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly created vertex \Vertex{\vidConstructionY} containing the variable reference \texttt{y}.
The resulting graph represents the expression \texttt{x * y}.

@@

TODO: delete "x" instead of "y"
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   {}
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 y [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Change-b}
TODO: Delete this figure
\newedge{ChangeZ}{R}
\newvertex{ChangeZ}{z}
\begin{tikzpicture}[framed]
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vChangeZ" [> "\eChangeZ"]
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{y}_{6}$"
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change.
\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{5}.
\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{8} and \Edge{7}.}
\label{fig:Change}
\end{figure}

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-0.5,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-b}
\newedge{WrapPlus}{Root}
\newvertex{WrapPlus}{+}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (-0.5cm,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-c}
\newedge{WrapTimes}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eWrapTimes"'] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping.
\autoref{fig:Wrapping-a}: \autoref{fig:Change-b} after deleting \Edge{1}.
\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{8} and \Edge{7}.
\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{9}.
}
\label{fig:Wrapping}
\end{figure}

@@

Next, Alice types~\texttt{+} when the cursor is on \Vertex{2}.
If this were a hole, this would simply put a vertex with the~\texttt{+} constructor
as a child of \Vertex{0}.
However, this position is already occupied by some code.

A naive editor could add the~\texttt{+} by creating a new
edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{+} and the existing \texttt{var}
as shown in Figure~REF:TODO.

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Move-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (0cm,-3cm) graph [graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.45\linewidth}
\centering
\caption{}
\label{fig:Move-b}
\newedge{MoveTimes}{R}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \Vertex{2} from the left child to the right child of \Vertex{8}.
\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{9}.
\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{11}.}
\label{fig:Move}
\end{figure}

@@

For example, consider \autoref{fig:Move}, which represents
the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
to its right child.
\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
First, \Edge{7} is deleted, which results in \autoref{fig:Move-a}
where the rooted code is now~\texttt{\hole{} + \hole{}},
and \Vertex{2} is orphaned.
Next \Edge{11} is added, which results in the bottom graph
where the rooted code is now~\texttt{\hole{} + 1 * 2},
and \Vertex{2} is no longer orphaned.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-a}
\newedge{DifferentAlice}{L}
\newvertex{DifferentAlice}{u}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-b}
\newedge{DifferentBob}{R}
\newvertex{DifferentBob}{v}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Different Parts of the Code.
\autoref{fig:Editing Different Parts of the Code-a}:~\autoref{fig:Move-b} after Alice replaces \Vertex{4} with \Vertex{12}.
\autoref{fig:Editing Different Parts of the Code-b}:~\autoref{fig:Move-b} after Bob replaces \Vertex{6} with \Vertex{14}.
\autoref{fig:Editing Different Parts of the Code-c}:~The
result of merging \autoref{fig:Editing Different Parts of the Code-a}
with \autoref{fig:Editing Different Parts of the Code-b}.}
\label{fig:Editing Different Parts of the Code}
\end{figure}

@@

Alice changes the~\texttt{x} to a~\texttt{u},
and Bob changes the~\texttt{z} to a~\texttt{v}.
Before transmitting their edits to each other,
Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
For Alice, this involves deleting \Edge{3} and creating \Edge{13}.
For Bob, this involves deleting \Edge{5} and creating \Edge{15}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-a}
\newedge{NestedAlice}{L}
\newvertex{NestedAlice}{w}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-b}
\newedge{NestedBob}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Nested Parts of the Code.
\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{14} with \Vertex{16}.
\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{2} to the left child of \Vertex{8}.
\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
\label{fig:Editing Nested Parts of the Code}
\end{figure}

>>>>

%% Title information
% \title[Grove]{Convergent Collaborative Structure Editing}
\title[Grove]{Grove: A Convergent Collaborative Structure-Editor Calculus}
%\subtitle{Subtitle}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we represent expressions with holes as graphs,
\autoref{sub:Single-User Actions} explores edits by a single user, Alice,
and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

TODO: move subcaptions into master caption

TODO: colors for different editors

TODO: label edges of transitions between graphs

TODO: Remove subfigure captions
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{A simple tree with a hole.
\autoref{fig:A simple tree with a hole-a}:~User interface.
\autoref{fig:A simple tree with a hole-b}:~Graph representation.}
\label{fig:A simple tree with a hole}
\end{figure}

@@

Program state is a directed graph with a distinguished root vertex.
%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

@@

Finally, holes are represented by the absence of a child.
For example, in \autoref{fig:A simple tree with a hole-b} the absence of a \texttt{right} edge coming from \Vertex{\vidSimpleTimes}
corresponds to the hole on the right of the~\texttt{*} in~\texttt{x * \hole}.

@@

\subsection{Single-User Actions}
\label{sub:Single-User Actions}

In the remainder of this section,
we will consider user actions, starting
in this subsection with single-user
actions and continuing in \autoref{sub:Multi-user Edits}
with multi-user actions.

Each user action that we consider corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

User actions act relative to a cursor. 
We do not model cursor sharing and the cursor does not appear in the underlying graph.
We discuss cursor representations and cursor sharing in more detail in \autoref{sub:Cursors}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\columnwidth}
\centering
\caption{}
\label{fig:Construction-a}
TODO: remove this figure
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\columnwidth}
\centering
\caption{}
\label{fig:Construction-b}
\newedge{ConstructionY}{R}%
\newvertex{ConstructionY}{y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vConstructionY" [> "\eConstructionY"]
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Construction.
\autoref{fig:Construction-a}:~User interface.
\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
\label{fig:Construction}
\end{figure}

\todo{"user actions" and "graph edits"}
To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
created vertex, \Vertex{\vidConstructionY}, containing the variable reference \texttt{y}.
The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.

@@

TODO: delete "x" instead of "y"

\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   {}
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 y [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Change-b}
TODO: Delete this figure
\newedge{ChangeZ}{R}
\newvertex{ChangeZ}{z}
\begin{tikzpicture}[framed]
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vChangeZ" [> "\eChangeZ"]
  }
 }
};
\path (1cm,-2cm) graph[graph style] {
 "$\texttt{y}_{6}$"
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Change.
\autoref{fig:Change-a}:~\autoref{fig:Construction-b} after deleting \Edge{\eidConstructionY}.
\autoref{fig:Change-b}:~\autoref{fig:Change-a} after adding \Vertex{\vidChangeZ} and \Edge{\eidChangeZ}.}
\label{fig:Change}
\end{figure}

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-0.5,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-b}
\newedge{WrapPlus}{Root}
\newvertex{WrapPlus}{+}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (-0.5cm,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-c}
\newedge{WrapTimes}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eWrapTimes"'] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping.
\autoref{fig:Wrapping-a}:~\autoref{fig:Change-b} after deleting \Edge{\eidSimpleTimes}.
\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{\vidWrapPlus} and \Edge{\eidWrapPlus}.
\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{\eidWrapTimes}.}
\label{fig:Wrapping}
\end{figure}

@@

TODO: revise: "user action" "graph edit"

Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
If this were a hole, this would cause a construction edit to be performed as in Sec.TODO above.
However, this position is already occupied by the \texttt{*} expression, 
so instead, typing~\texttt{+} performs a wrapping edit.
This corresponds to the following sequence of graph actions, shown in \autoref{fig:Wrapping}: 
(a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex,
(c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
(The choice of wrapping with a bias for the left child position is arbitrary.)

% A naive editor could add the~\texttt{+} as a new child of the root.
% However,

by creating a new
edge from the \texttt{L} index of the \texttt{*} to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{+} and the existing \texttt{var}
as shown in Figure~REF:TODO.

You can have a second edge,
we discuss conflicts in section TODO
In the single-user setting, we don't want to create conflicts, so instead we perform a wrapping action.

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Move-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (0cm,-3cm) graph [graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.45\linewidth}
\centering
\caption{}
\label{fig:Move-b}
\newedge{MoveTimes}{R}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.
\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{\eidWrapTimes}.
\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{\eidMoveTimes}.}
\label{fig:Move}
\end{figure}

@@

For example, consider \autoref{fig:Move}, which represents
the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
to its right child.
\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
First, \Edge{\eidWrapPlus} is deleted, which results in \autoref{fig:Move-a}
where the rooted code is now~\texttt{\hole{} + \hole{}},
and \Vertex{\vidSimpleTimes} is orphaned.
Next \Edge{\eidMoveTimes} is added, which results in the bottom graph
where the rooted code is now~\texttt{\hole{} + 1 * 2},
and \Vertex{\vidSimpleTimes} is no longer orphaned.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-a}
\newedge{DifferentAlice}{L}
\newvertex{DifferentAlice}{u}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-b}
\newedge{DifferentBob}{R}
\newvertex{DifferentBob}{v}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Different Parts of the Code.
\textbf{(a)}:~\autoref{fig:Move-b} after Alice replaces \Vertex{\vidSimpleX} with \Vertex{\vidDifferentAlice}.
\textbf{(b)}:~\autoref{fig:Move-b} after Bob replaces \Vertex{\vidConstructionY} with \Vertex{\vidDifferentBob}.
\textbf{(c)}:~The
result of merging \autoref{fig:Editing Different Parts of the Code-a}
with \autoref{fig:Editing Different Parts of the Code-b}.}
\label{fig:Editing Different Parts of the Code}
\end{figure}

@@

Alice changes the~\texttt{x} to a~\texttt{u},
and Bob changes the~\texttt{z} to a~\texttt{v}.
Before transmitting their edits to each other,
Alice thus has the graph in \autoref{fig:Editing Different Parts of the Code-a}
and Bob has the graph in \autoref{fig:Editing Different Parts of the Code-b}.
For Alice, this involves deleting \Edge{3} and creating \Edge{13}.
For Bob, this involves deleting \Edge{\eidConstructionY} and creating \Edge{15}.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-a}
\newedge{NestedAlice}{L}
\newvertex{NestedAlice}{w}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-b}
\newedge{NestedBob}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Nested Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eNestedBob"'] -> {
    "\vNestedAlice" [> "\eNestedAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Nested Parts of the Code.
\autoref{fig:Editing Nested Parts of the Code-a}:~\autoref{fig:Editing Nested Parts of the Code-c} after Alice replaces \Vertex{\vidDifferentAlice} with \Vertex{\vidNestedAlice}.
\autoref{fig:Editing Nested Parts of the Code-b}:~\autoref{fig:Editing Nested Parts of the Code-c} after Bob move \Vertex{\vidSimpleTimes} to the left child of \Vertex{\vidWrapPlus}.
\autoref{fig:Editing Nested Parts of the Code-c}:~The result of merging \autoref{fig:Editing Nested Parts of the Code-a} with \autoref{fig:Editing Nested Parts of the Code-b}.}
\label{fig:Editing Nested Parts of the Code}
\end{figure}
